\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-02-20 paper.qmd
Added cluster label from result 1 and result 2, side by side 2D layout coloured by labels in result1, result2

>>>>

# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?
```
<!--Generate true label for PBMC3k-->

UMAP_pbmc_with_result1_label <- UMAP_pbmc |>
  dplyr::mutate(result1_class_label = case_when(
    ((UMAP1 >=-12) & (UMAP1 <= -5) & (UMAP2 >= 6) & (UMAP2 <= 18)) ~ "cluster 1",
    ((UMAP1 >-5) & (UMAP1 <= 7) & (UMAP2 >= 4.8) & (UMAP2 <= max(UMAP2))) ~ "cluster 2", ((UMAP1 >=6) & (UMAP1 <= 15) & (UMAP2 >= -4) & (UMAP2 < 4.8)) ~ "cluster 3",
    ((UMAP1 >=-3) & (UMAP1 <= 6.6) & (UMAP2 >= min(UMAP2)) & (UMAP2 <= 4.8)) ~ "cluster 4",
    ((UMAP1 >=min(UMAP1))& (UMAP1 < -3) & (UMAP2 >= -8) & (UMAP2 <= -1)) ~ "cluster 5",
    ((UMAP1 >=min(UMAP1)) & (UMAP1 <= -7) & (UMAP2 >= -3) & (UMAP2 <= 3)) ~ "cluster 6",
    .default = "no class"
    ))

### Reassign some labels

UMAP_pbmc_with_result1_label <- UMAP_pbmc_with_result1_label |>
  dplyr::mutate(result1_class_label = if_else((ID %in% c(2619, 1158)), "cluster 3", result1_class_label)) |>
  dplyr::mutate(result1_class_label = if_else((ID %in% c(948, 285, 158, 268, 731, 1152, 1372, 1651, 2010, 150, 2304, 2546, 487, 2176, 2157, 2451, 264, 63, 1234, 947, 1126, 1699, 1232, 1492, 1488, 2040, 2550, 1313, 361, 372, 1515, 1264, 1307, 1726, 517, 2568, 2274, 591, 1703, 1097, 430, 1556, 2133, 1232, 1194, 1888, 1796, 494, 1194)), "cluster 6", result1_class_label))

## Import detour results
detour_results8 <- read_csv("data/pbmc/pbmc_3k_festem/detourr_export8.csv")

### Rename labels

detour_results8 <- detour_results8 |>
  dplyr::mutate(result2_class_label = case_when(
    colour == "000000" ~ "cluster 4",
    colour == "49df20" ~ "cluster 2",
    colour == "619cff" ~ "cluster 3",
    colour == "d82cd3" ~ "cluster 1",
    colour == "df2020" ~ "cluster 5",
    colour == "ffbd61" ~ "cluster 6",
    .default = "no cluster"

  ))

detour_results8 <- detour_results8 |>
  dplyr::mutate(ID = training_data_pbmc$ID)

df_res_1_2_label <- inner_join(UMAP_pbmc_with_result1_label, detour_results8, by = c("ID" = "ID"))

result1_2d_layout <- df_res_1_2_label |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = result1_class_label))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

result2_2d_layout <- df_res_1_2_label |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = result2_class_label))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#d82cd3", "#49df20", "#619cff", "#000000", "#df2020", "#ffbd61", "yellow")) +
  # geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

table(df_res_1_2_label$result1_class_label, df_res_1_2_label$result2_class_label)

result1_2d_layout + result2_2d_layout
```

### 2024-02-20 paper.tex
Added cluster label from result 1 and result 2, side by side 2D layout coloured by labels in result1, result2

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, breakable, interior hidden, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

\begin{verbatim}

            cluster 1 cluster 2 cluster 3 cluster 4 cluster 5 cluster 6
  cluster 1       138        13         0         0         0         0
  cluster 2        11       509         0         1         0         0
  cluster 3         1         0       348         1         0         1
  cluster 4         0         0         3      1029        64         2
  cluster 5         0         0         0        85       226        32
  cluster 6         0         0         0         0        19       139
\end{verbatim}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-71-1.pdf}
